All training and evaluation runs reported here were carried out on Kaggle's free dual-T4 environment (Tesla T4 $\times$ 2, 15~GiB device memory each) with Python 3.12 and Ultralytics 8.4.46~\cite{jocher2023ultralytics}. Joint pretraining took roughly 75--85 wall-clock minutes; each fine-tuning run took 0.4--5 minutes depending on budget and freeze setting. The full experimental matrix --- 4 methods $\times$ 5 budgets $\times$ 3 seeds, minus a few historical duplicates introduced when an earlier label-budget configuration capped to the same effective sample count --- comprises 72 fine-tune runs in total.

Splits are the Roboflow-export defaults for each crop. We did not modify the validation or test splits, which keeps our test numbers comparable to other work using the same public datasets. The kiwi test set on which all reported numbers are computed contains 61 images and 369 flower instances. We use Ultralytics' \texttt{val(split="test")} call so that the test mAP@0.5 and mAP@0.5{:}0.95 numbers in the results section match what a downstream user would measure with the same script and a checkpoint we publish.

Three random seeds (42, 43, 44) drive both the per-budget subsample of the kiwi training images and the model initialization where applicable. We chose three seeds rather than five or ten for compute reasons; in practice, three seeds proved enough to expose two failure modes (the from-scratch variance at very low budgets and the unfrozen-CrossPoll collapse at $n{=}25$) but they are at the lower end of statistical reliability, and we report ranges rather than confidence intervals throughout.

Mean Average Precision at IoU=0.5 (mAP@0.5) is the headline metric. We additionally report mAP@0.5{:}0.95 where the underlying run logged it; some early baseline runs predate the addition of the secondary metric to our logging script and therefore have empty cells in the secondary table. Run-level results are persisted in \texttt{results/summary.csv}, and the table and figure in this paper are generated from that file by \texttt{scripts/build\_paper\_assets.py}, so the manuscript and the data cannot drift.
