\label{sec:intro}

\subsection{Motivation and Problem Statement}

Global pollinator populations are declining at alarming rates due to habitat loss, pesticide exposure, and climate change \cite{vanEngelsdorp2010}. This decline threatens agricultural productivity, with an estimated 75\% of global crops dependent on animal pollinators for reproduction or yield \cite{Klein2007}. In response, autonomous robotic pollination has emerged as a promising solution to maintain crop yields in pollinator-limited environments. However, robotic systems face a critical challenge: they must rapidly detect flowers across diverse crop species—each with distinct visual morphologies, growth patterns, and environmental conditions—without requiring months of manual annotation for each new crop.

Current approaches to flower detection in robotic systems fall into two categories, both with significant limitations:

\begin{enumerate}
    \item \textbf{Crop-specific training}: Training separate detectors for each crop requires 1,000--5,000 manually-annotated flower bounding boxes per crop \cite{GeneMola2021}. At 50--100 annotations per hour, this translates to 10--100 hours of labeling per crop—prohibitively expensive for growers deploying systems to multiple varieties.
    
    \item \textbf{Generic transfer learning}: Fine-tuning COCO-pretrained detectors (e.g., YOLOv8) directly on small flower datasets typically yields 15--30\% lower mAP than crop-trained models due to the large visual domain gap between natural objects (COCO) and agricultural flowers \cite{Zhang2022}.
\end{enumerate}

This creates a critical bottleneck: deploying robotic pollinators economically requires methods that achieve high accuracy with minimal crop-specific labels while exploiting available flower datasets from related crops.

\subsection{Core Hypothesis and Research Question}

We hypothesize that \textit{multi-source joint pretraining}—where a detector learns flower representations from multiple diverse crops simultaneously—provides richer, more generalizable features than single-source transfer, enabling faster and more accurate adaptation to new crops with fewer labels.

\textbf{Research Question}: Can a detector jointly pretrained on flowers from multiple source crops adapt to unseen target crops using significantly fewer labeled images than COCO-pretrained baselines, while remaining competitive with crop-specific training?

\subsection{Contributions}

This work makes three primary contributions:

\begin{enumerate}
    \item \textbf{CrossPoll Framework}: We propose a practical, multi-crop transfer learning pipeline specifically designed for flower detection. The framework jointly pretrains YOLOv8 on three source crops, then fine-tunes with varying label budgets (25--500 images) on a held-out target crop. This approach is architecture-agnostic and easily adaptable to new crop combinations.
    
    \item \textbf{Systematic Empirical Evaluation}: We conduct an extensive evaluation across three baseline methods, five label budgets, and three random seeds, demonstrating that multi-source transfer provides consistent 15--25\% mAP gains over single-source COCO transfer at all budget levels. Our results establish clear evidence that multi-crop pretraining is more effective than generic pretrained models for agricultural detection.
    
    \item \textbf{OrchardFlow Dataset}: We curate and release OrchardFlow, a multi-crop flower dataset with 16,613 images across four crops (Kiwi, Apple, Strawberry, Tomato) in YOLO format. This resource enables reproducible evaluation of transfer learning methods in agricultural contexts and reduces barriers to future research in this domain.
\end{enumerate}

\subsection{Impact and Broader Significance}

Sample-efficient flower detection accelerates robotic pollination deployment. By reducing per-crop labeling from hundreds of hours to tens of hours, CrossPoll enables cost-effective multi-crop pollination systems that can adapt to new varieties and cultivars rapidly. Beyond agriculture, the multi-source transfer approach generalizes to other fine-grained detection tasks with limited data, including industrial quality control, medical imaging, and ecological monitoring.

\subsection{Paper Organization}

The remainder of this paper is organized as follows: Section~\ref{sec:related} reviews transfer learning, object detection, domain adaptation, and agricultural computer vision literature. Section~\ref{sec:methods} describes the CrossPoll pipeline, dataset preparation, and experimental design. Section~\ref{sec:experiments} details our experimental setup, including baselines, metrics, and evaluation protocol. Section~\ref{sec:results} presents quantitative results and comparative analysis. Section~\ref{sec:discussion} discusses findings, limitations, and implications for robotic systems. Finally, Section~\ref{sec:conclusion} concludes with future work.
