CrossPoll has two stages. We first pretrain a YOLOv8n detector~\cite{jocher2023ultralytics} on a label-harmonized union of three source crops (apple, strawberry, tomato), producing a flower-specialized feature extractor. We then fine-tune that extractor on the held-out target crop (kiwi) at five label budgets, with and without backbone freezing. Both stages reuse the same architecture and training stack so that any difference in test-time accuracy is attributable to the transfer protocol rather than to model capacity.

\subsection{Datasets and label harmonization}

The four source datasets come from Roboflow Universe~\cite{roboflow2024} as YOLO-format exports with predefined train/val/test splits (roughly 70/20/10 per crop). Apple has a single \emph{flower} class with 10{,}379 images. Strawberry contains 5{,}419 images annotated for three classes: flower, ripe fruit, and unripe fruit. Tomato is the smallest source set with 513 images and three generic class labels. Kiwi is the held-out target with 302 images and one flower class; the kiwi train split contains only 181 images, which sets the upper bound on our few-shot budget sweep.

For joint pretraining we collapse all source-crop annotations to a single \texttt{flower} class. The mapping is conservative: apple's flower class is kept; strawberry's ripe and unripe fruit classes are dropped (they are not flowers); and all three tomato classes are mapped to flower. The tomato choice rests on the assumption, drawn from manual inspection of a small sample, that the three classes correspond to flower morphologies at different stages rather than to non-flower objects. We document this in the harmonization script (\texttt{scripts/harmonize\_labels.py}) so that a different mapping is one edit away. After harmonization the joint training set covers approximately 11{,}300 images of flowers across the three source crops.

We perform the harmonization by writing fresh YOLO label files into a parallel directory tree (\texttt{data/joint/<crop>/<split>}) while copying the corresponding image directories. We initially used symbolic links for the images but found that Ultralytics resolves the link before deriving the label path, which silently bypassed our harmonized labels and read the originals. The bug surfaced as a wave of \texttt{ignoring corrupt image/label} warnings in the log; copying the images fixed it.

\subsection{Joint pretraining}

The joint dataset is described to Ultralytics by a single YAML file that lists three train and three validation image directories with \texttt{nc=1, names=[flower]}. We initialize from the public COCO-pretrained \texttt{yolov8n.pt} weights rather than from scratch so that the network already has a sensible low-level filter bank. Training uses AdamW~\cite{loshchilov2019adamw} with an initial learning rate of $10^{-3}$, a cosine schedule~\cite{loshchilov2017sgdr}, batch size 16, $640 \times 640$ input resolution, and the default Ultralytics augmentation recipe (mosaic, HSV jitter, horizontal flip). We allow up to 150 epochs but stop early when the validation mAP@0.5 has not improved for 15 consecutive epochs. On a Kaggle dual-T4 environment the run converges in roughly 75--85 minutes and produces a checkpoint with mAP@0.5 of 0.685 and mAP@0.5{:}0.95 of 0.473 on the held-out joint validation split.

\subsection{Few-shot fine-tuning on the target crop}

For each of the five label budgets $n \in \{10, 25, 50, 100, 181\}$, three random seeds (42, 43, 44), and four method choices we run an independent fine-tune. The training subset is sampled deterministically from the kiwi train split using the seed, with image-label pairs symlinked into a per-run scratch directory and described to Ultralytics by a freshly written \texttt{data.yaml}. The validation split (\emph{kiwi/valid}) provides early-stopping signal; the test split (\emph{kiwi/test}, 61 images, 369 instances) provides the reported metrics.

The four methods differ only in initialization and in whether the backbone is frozen:
\begin{description}
  \item[Scratch.] Random YOLOv8n weights from the architecture YAML. No transfer.
  \item[COCO transfer.] Initialized from the public \texttt{yolov8n.pt} (COCO).
  \item[CrossPoll (full FT).] Initialized from our joint pretrain checkpoint. All weights are updated.
  \item[CrossPoll + freeze.] Initialized from the joint pretrain checkpoint, with the YOLOv8n backbone (modules 0--9 in the Ultralytics convention~\cite{gandhi2025fine}) held fixed. The neck (FPN~\cite{lin2017fpn}) and the detection head are trained.
\end{description}

The frozen variant is motivated by the observation that the joint pretrain produces highly domain-specialized features. Earlier work on YOLOv8 fine-tuning for fruit detection~\cite{gandhi2025fine} reports that freezing the first ten modules retains the pretrained representation while allowing task-specific adaptation in the head, and we found the same intuition to hold here.

Fine-tuning hyperparameters are shared across the four methods and across budgets: AdamW, cosine learning rate schedule, batch size 16, 640$\times$640 inputs, automatic mixed precision, and 150-epoch ceiling. The initial learning rate is $10^{-4}$, an order of magnitude below the joint pretrain rate; this is appropriate when starting from a near-converged checkpoint. Patience is 15 epochs for the unfrozen runs and 30 epochs for the frozen variant. We extended patience for the frozen runs because the head needs more updates to adapt when the backbone cannot move, and the early-epoch validation curves are noisier on small budgets.

The per-budget actual training sample count is capped at 181 (the kiwi train size), so a request for $n{=}500$ collapses to 181 images. We picked the budgets above the cap so the curve is anchored both at the very few-shot end ($n{=}10$) and at the realistic upper bound. All run identifiers, hyperparameter dumps, and per-epoch metrics are recorded by the training script and stored under \texttt{runs/} for reproducibility.
